---
title: Download Loop
linktitle: Download Loop
toc: false
type: book
date: "2021-01-25T00:00:00Z"
draft: false
# Add menu entry to sidebar.
# - name: Declare this menu item as a parent with ID `name`.
# - weight: Position of link in menu.
menu:
  rcourse:
    parent: Lesson 1
    weight: 60

weight: 60
---

```{r, include=FALSE}
local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    x <- xfun::split_lines(x)
    n <- nchar(x, 'width')
    i <- seq_along(x)  # line numbers
    n <- n + nchar(i)
    s <- knitr:::v_spaces(max(n) - n)
    x <- paste(x, s, '  # ', i, sep = '', collapse = '\n')
    hook_source(x, options)
  })
})
```


```{r eval=FALSE}
require(stringr)
require(readxl)
county_data<-data.frame()
years<-c(90:99, 0:19)
years<-str_pad(as.character(years), 2, "left", "0")
years_l<-length(years)
for (i in 1:years_l){
url<-paste0("https://www.bls.gov/lau/laucnty", years[i], ".xlsx")
destination<-paste0(workingdir, "/raw_data/bls_unemp_", years[i], ".xlsx")
download.file(url, destination, mode="wb")
temp_df <- read_excel(paste0(workingdir, "/raw_data/bls_unemp_", years[i],
  ".xlsx"),col_names=FALSE, skip=5)
temp_df<-temp_df[,-6]
colnames(temp_df)<-c("LAUS_code", "State_fips", "County_fips", "County_name",
  "Year", "Labor_force", "Employed", "Unemployed", "Unemp_rate")
county_data<-rbind(county_data, temp_df)
}
save(county_data, file=paste0(workingdir, "/processed_data/county_data_all"))
```

First let's think what we are trying to achieve. The links for the downloads are all in the form of https://www.bls.gov/lau/laucntyZZ.xlsx, where ZZ is two digits representing the year. These ZZ values run from "90"to "99" for years 1990-1999 and "00" to "19" for years 2000-2019.

Let's work through the above code line by line:

- Lines #1 & #2: load required packages.
- Line #3: Declare county_data will be a data.frame. Right now it is empty, but we will add to it.
- Line #4: Define a vector with elements 90-99 and 0-19. (Which will correspond to the years that we will pull)
- Line #5 : We have a vector for years, however if we notice in the url names we need this vector to include a leading 0 in front of the single 'character' digits (ie "01" instead of "1"). But we don't want a leading 0 in front of the double 'character' digits (ie We DON'T want "090"). Go back to the [String Manipulation](http://jmwestenberg.github.io/courses/rcourse/lesson1_03/) section if you need to refresh on this.
- Line #6 : Calculate the length of years and save as years_l
- Line #7 : See [String Manipulation](http://jmwestenberg.github.io/courses/rcourse/lesson1_02/) if defining the for loop does not make sense.
- Line #8 : We are creating the character string for the url for the download link. Since they all take the form of https://www.bls.gov/lau/laucntyZZ.xlsx, we can use one element of our years vector at a time. (See [String Manipulation](http://jmwestenberg.github.io/courses/rcourse/lesson1_03/) for explanation on paste0)
- Line #9 : This is of similar spirit to line #4, but this is defining the path/filename of the excel file we will save. 
- Line #10 : This line is just telling R to download the file at that url, save it to the defined location/name, and to read it as a non-raw text form. (see [Downloading/Reading Data](http://jmwestenberg.github.io/courses/rcourse/lesson1_04/) if unclear.)
- Lines #11-#15: See [Downloading/Reading Data](http://jmwestenberg.github.io/courses/rcourse/lesson1_04/) for a direct explanation.
- Line #16: rbind appends data. Hence since all of our data has the same format and has a variable indicating the year, we can simply append.
- Line #18: save the combined data to our processed_data folder as a Rda type.

We now have our data cleaned and saved for our next lesson when we will start to work with it more!