---
title: Writing Loops
linktitle: Writing Loops
toc: false
type: book
date: "2021-02-05T00:00:00Z"
draft: false
# Add menu entry to sidebar.
# - name: Declare this menu item as a parent with ID `name`.
# - weight: Position of link in menu.
menu:
  rcourse:
    parent: Lesson 1
    weight: 20

weight: 20
---

```{r, include=FALSE}
local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    x <- xfun::split_lines(x)
    n <- nchar(x, 'width')
    i <- seq_along(x)  # line numbers
    n <- n + nchar(i)
    s <- knitr:::v_spaces(max(n) - n)
    x <- paste(x, s, '  # ', i, sep = '', collapse = '\n')
    hook_source(x, options)
  })
})
```

We will be using [county level employment/labor force data from BLS](https://www.bls.gov/lau/home.htm). We will be using the **Labor force data by county, yearly annual averages**. There is data from 1990-2019 (as of writing these notes). To start we are going to download this data and then read it into R.

We could download these all one by one, or we could write a simple loop to read them in for us. To do this let's first look at a very simple loop structure:

```{r class.source = c("numCode", "R", "numberLines")}
letters<-c("a", "b", "c", "d", "e")
letters_l<-length(letters)
for (i in 1:letters_l){
  print(letters[i])
}
```
Going line by line:

- Line 1: Define a vector of letters.
- Line 2: Report the number of elements in our letter vector and save as letters_l
- Line 3: Defining for loop. Our index will be i, and it will run from 0 to however long letters vector is (try adding some more letters!) Note our for loop action is defined within the curly braces.
- Line 4: For every i defined in Line 3 we want to print the corresponding element in the vector letters.

Okay, great. How does this help us with reading in the data? We'll get to that in the next section.
