---
title: Writing Loops
linktitle: Writing Loops
toc: false
type: docs  # Do not modify.
date: "2021-01-25T00:00:00Z"
draft: false
# Add menu entry to sidebar.
# - name: Declare this menu item as a parent with ID `name`.
# - weight: Position of link in menu.
menu:
  rcourse:
    parent: Lesson 1
    weight: 1

weight: 20
---

```{r, include=FALSE}
local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    x <- xfun::split_lines(x)
    n <- nchar(x, 'width')
    i <- seq_along(x)  # line numbers
    n <- n + nchar(i)
    s <- knitr:::v_spaces(max(n) - n)
    x <- paste(x, s, '  # ', i, sep = '', collapse = '\n')
    hook_source(x, options)
  })
})
```

We will be using [county level employment/labor force data from BLS](https://www.bls.gov/lau/home.htm). We will be using the **Labor force data by county, yearly annual averages**. There is data from 1990-2019 (as of writing these notes). To start we are going to download this data and then read it into R.

We could download these all one by one, or we could write a simple loop to read them in for us. To do this let's first look at a very simple loop structure:

```{r class.source = c("numCode", "R", "numberLines")}
i=1
repeat{
print(i)
if (i==5){break}
i<-i+1
}
```
Going line by line:
1. We make i the numeric value we will be looping over starting at 1. 
2. Repeat the action in the curly braces (note be careful, without an exit this will continuously loop)
3. Print out the current value that i is taking.
4. The break for our loop, when i takes the value 5, exit the loop
5. If the line 4 break statement is not true, then increment i by 1.

Taken all together we are printing then incrementing i by one while i is less than or equal to 5.

Okay, great. How does this help us with reading in the data? We'll get to that in the next section.
