---
title: Concatenating Strings
linktitle: Concatenating Strings
toc: false
type: docs  # Do not modify.
date: "2021-01-25T00:00:00Z"
draft: false
# Add menu entry to sidebar.
# - name: Declare this menu item as a parent with ID `name`.
# - weight: Position of link in menu.
menu:
  rcourse:
    parent: Lesson 1
    weight: 2

weight: 40
---

```{r, include=FALSE}
local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    x <- xfun::split_lines(x)
    n <- nchar(x, 'width')
    i <- seq_along(x)  # line numbers
    n <- n + nchar(i)
    s <- knitr:::v_spaces(max(n) - n)
    x <- paste(x, s, '  # ', i, sep = '', collapse = '\n')
    hook_source(x, options)
  })
})
```


Before we get into downloading our data. Let's learn how to concatenate strings in R.

```{r}
a<-"This is a the start"
b<-"of a sentence"
print(paste0(a,b))
print(paste(a,b))
```
Notice we define a and b as strings. What paste and paste0 do are combine these strings into one string. We can see that paste places a space between the two strings while paste0 does not. paste0 comes in quite handy for working with file pathing as we will see. Yes it's that easy! 

## Note/Extras

We will not be using this for this project but it may be useful to know we can concatenate two vectors of strings as well. 
```{r}
a<-c("This is a the start", "Now we have", "This really is", "Economics is")
b<-c("of a sentence.", "another sentence.", "quite handy.", "awesome!")
print(paste(a,b))
```

Another command I have used a lot within the 