---
title: group_by and summarise
linktitle: group_by and summarise
toc: false
type: book
date: "2021-02-05T00:00:00Z"
draft: false
# Add menu entry to sidebar.
# - name: Declare this menu item as a parent with ID `name`.
# - weight: Position of link in menu.
menu:
  rcourse:
    parent: Lesson 2
    weight: 30

weight: 30
---


```{r, include=FALSE}
local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    x <- xfun::split_lines(x)
    n <- nchar(x, 'width')
    i <- seq_along(x)  # line numbers
    n <- n + nchar(i)
    s <- knitr:::v_spaces(max(n) - n)
    x <- paste(x, s, '  # ', i, sep = '', collapse = '\n')
    hook_source(x, options)
  })
})

require(tidyverse)
require(xtable)
require(knitr)
require(kableExtra)

workingdir<-"C:/Users/weste/Documents/GitHub/Rcourse/"
source(paste0(workingdir, "workingdir.r"))

load(paste0(workingdir, file.path("processed_data", "unemp_data.rda")))

indiana<-unemp_data %>% filter(State_fips=="18")
```


Our next chunk of code gets a bit more complex, let's take it peice by peice though.

```{r eval=FALSE}
  indiana %>% group_by(Year) %>% 
    summarise(Average=mean(Labor_force))
```

Let's again take it from the end and work our ways backwards through it. The 3rd line above is calculating a new variable within a group that is called 'Average'. The group (line 2) that we are calculating within is each Year for our Indiana data.frame. That is this snippet of code will give us one value back for each year, the average Labor Force level across Indiana's counties. 

```{r echo=FALSE}
temp<-
  indiana %>% group_by(Year) %>% 
    summarise(Average=mean(Labor_force))
scroll_box(kable(temp), height="300px", width="180px")
```

This group_by command paired with summarise is quite useful. For example going back to our original data.frame, county_data. Suppose we wanted to do what we just did but for all states. That is calculate within each state and year combination. With the group_by command this is quite simple! (I print the first 100 lines of output below)

```{r eval=FALSE}
unemp_data %>% 
  group_by(Year, State_fips) %>% 
  summarise(Average=mean(Labor_force))
```

```{r echo=FALSE}
all_ave<-unemp_data %>% group_by(Year, State_fips) %>% summarise(Average=mean(Labor_force))
scroll_box(kable(all_ave[1:100,]), height="300px", width="290px")
```

Average is just one statistic we may want to calculate, how about percentiles, minimum, maximum, standard deviation. our summarise command will accept more than one argument, we just have to separate by a comma. Let's also save this as a new data.frame for use in future lessons.

```{r}
indiana_laborforce<-
  indiana %>% group_by(Year) %>% 
  summarise(Min=min(Labor_force),
    p10th=quantile(Labor_force, c(0.1)),
    p25th=quantile(Labor_force, c(0.25)),
    p50th=quantile(Labor_force, c(0.5)),
    p75th=quantile(Labor_force, c(0.75)),
    p90th=quantile(Labor_force, c(0.9)),
    Max=max(Labor_force),
    Average=mean(Labor_force),
    StDev=sd(Labor_force))
```

```{r echo=FALSE}
scroll_box(kable(indiana_laborforce), height="300px")
```