---
title: select
linktitle: select
toc: false
type: book
date: "2021-02-05T00:00:00Z"
draft: false
# Add menu entry to sidebar.
# - name: Declare this menu item as a parent with ID `name`.
# - weight: Position of link in menu.
menu:
  rcourse:
    parent: Lesson 2
    weight: 50

weight: 50
---

```{r, include=FALSE}
local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    x <- xfun::split_lines(x)
    n <- nchar(x, 'width')
    i <- seq_along(x)  # line numbers
    n <- n + nchar(i)
    s <- knitr:::v_spaces(max(n) - n)
    x <- paste(x, s, '  # ', i, sep = '', collapse = '\n')
    hook_source(x, options)
  })
})

require(tidyverse)
require(xtable)
require(knitr)
require(kableExtra)

workingdir<-"C:/Users/weste/Documents/GitHub/Rcourse/"
load(paste0(workingdir, file.path("processed_data", "unemp_data.rda")))

indiana<-unemp_data %>% filter(State_fips=="18")

indiana<-
  indiana %>% 
  mutate(unemp_rate_calc=round((Unemployed/Labor_force)*100, digits=1))
```


How can we go about dropping or keeping certain variables?

Say we wanted to drop the unemp_rate_calc, Labor_force, and Employed?
```{r}
temp<-
  indiana %>% 
  select(-c(unemp_rate_calc, Labor_force,  Employed))
head(temp)
```

Now what if we wanted to keep just State_fips, County_fips, Year, and Unemp_rate?

```{r}
temp<-
  indiana %>% 
  select(c(State_fips, County_fips, Year, Unemp_rate))
head(temp)
```

That is if we include the "-" before our variables, we are telling dplyr through our select argument to drop the variables we list. While if we don't have the "-", we are telling dplyr to keep only these variables.
