---
title: mutate
linktitle: mutate
toc: false
type: book
date: "2021-02-05T00:00:00Z"
draft: false
# Add menu entry to sidebar.
# - name: Declare this menu item as a parent with ID `name`.
# - weight: Position of link in menu.
menu:
  rcourse:
    parent: Lesson 2
    weight: 40

weight: 40
---

```{r, include=FALSE}
local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    x <- xfun::split_lines(x)
    n <- nchar(x, 'width')
    i <- seq_along(x)  # line numbers
    n <- n + nchar(i)
    s <- knitr:::v_spaces(max(n) - n)
    x <- paste(x, s, '  # ', i, sep = '', collapse = '\n')
    hook_source(x, options)
  })
})

require(tidyverse)
require(xtable)
require(knitr)
require(kableExtra)

workingdir<-"C:/Users/weste/Documents/GitHub/Rcourse/"
load(paste0(workingdir, file.path("processed_data", "unemp_data.rda")))

indiana<-unemp_data %>% filter(State_fips=="18")
```


How about if we don't want to calculate summary statistics within a group, but just want to calculate a new variable from each observation? Consider if we had unemployed and labor force levels but did not have unemployment rate, how could we go about calculating it with mutate?

```{r}
indiana<-
  indiana %>% 
  mutate(unemp_rate_calc=round((Unemployed/Labor_force)*100, digits=1))
```
Note since we are defining our new data frame as indiana, the one we are manipulating, we are in this case overwriting the indiana data.frame. Since the unemployment rate in our file was listed as percent and rounded to the nearest tenth, I did the same for our calculated. (digits=1 means one decimal place)

Now let's compare the first 20 rows of the given unemployment rate with the one we just calculated.

```{r echo=FALSE}
scroll_box(kable(indiana[1:20, 9:10]), height="300px", width="280px")
```