---
# Page metadata.
title: "Lesson 0"
date: "2021-01-24T00:00:00Z"
lastmod: "2021-01-24T00:00:00Z"
draft: false  # Is this a draft? true/false
type: docs  # Do not modify.

# Add menu entry to sidebar.
# - name: Declare this menu item as a parent with ID `name`.
# - weight: Position of link in menu.
menu:
  intror:
    name: "Lesson 0"
    weight: 2
---
```{r, include=FALSE}
local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    x <- xfun::split_lines(x)
    n <- nchar(x, 'width')
    i <- seq_along(x)  # line numbers
    n <- n + nchar(i)
    s <- knitr:::v_spaces(max(n) - n)
    x <- paste(x, s, '  # ', i, sep = '', collapse = '\n')
    hook_source(x, options)
  })
})
```


## Getting Started in R

I assume you have R installed and running. There are plenty of guides online on how to do this. 


Let's first define some arrays within R.

These can be numeric based, in this case integer.
```{r class.source = c("numCode", "R", "numberLines")}
A<-c(5,6,7)
B<-c(7,8,9)
```
Let's dig into what is going on a bit more. We are telling R to define a vector, this is the c( ) part, with elements 1,2,3 and give that vector a name A. The backwards arrow tells R what is the name and what is the element we are defining. 

We can make character based vectors as well.

```{r class.source = c("numCode", "R", "numberLines")}
C<-c("a", "b", "c")
D<-c("d", "e", "f")
```

We can then combine these vectors into a dataframe (this is relevant for when we start thinking about reading in/manipulating actual data). Since all of our vectors are length three, we can easily create a dataframe (think a matrix) where our column names will be the name of the vectors, and the rows will be the elements of the vectors.

```{r class.source = c("numCode", "R", "numberLines")}
first_dataframe<-data.frame(A,B,C,D)
print(first_dataframe)
```

Now first_dataframe is going to be of similar format as we will typically have when we read in data from excel files into R.

```{r class.source = c("numCode", "R", "numberLines")}
print(first_dataframe[1,])
```